Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-10-05 PAPER.tex
Fixed algebra mistake in Appendix

\Chat^{-1}\x = \frac{\C^{-1}\x}{1-\eta} \left[ 1 - \frac{\eta g}{1+ \eta (g-1)}\right],

@@

\hat{F}_{\alpha \alpha} = \frac{F_{\alpha \alpha}}{(1-\eta)^2} \left[ 1 -\frac{\eta g}{1+ \eta (g-1)} \frac{h_{\alpha \alpha}}{F_{\alpha \alpha}} + \frac{1}{2} \left( \frac{\eta g}{1+ \eta (g-1)} \right)^2 \frac{h_\alpha^2}{F_{\alpha \alpha}}\right],

@@

\hat{p}_\alpha^\textrm{loss} &=& \frac{1}{2} \frac{h_\alpha}{F_{\alpha \alpha}} \left[ 1 - \frac{\eta g}{1+ \eta (g-1)}\right]^2  \left[ 1 -\frac{\eta g}{1+ \eta (g-1)} \frac{h_{\alpha \alpha}}{F_{\alpha \alpha}} + \frac{1}{2} \left( \frac{\eta g}{1+ \eta (g-1)} \right)^2 \frac{h_\alpha^2}{F_{\alpha \alpha}}\right]^{-1} \nonumber \\
&=& \frac{1}{2} \frac{h_\alpha}{F_{\alpha \alpha}}\left(1-2\beta + \beta^2\right)\sum_{n=0}^\infty \sum_{k=0}^n {n \choose k} \left( \frac{\beta h_{\alpha \alpha}}{F_{\alpha \alpha}}\right)^k \left( - \frac{\beta^2 h_\alpha^2}{F_{\alpha \alpha}}\right)^{n-k},

@@

where $\beta \equiv \eta g/[1+ \eta (g-1)]$. To proceed, we take the ensemble average of $\hat{p}_\alpha^\textrm{loss}$ to understand its average statistical properties. Inspecting Eq. \eqref{eq:phatlossexpanded}, we see that its expectation value is a linear sum of terms of the form $\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle$, where $a$, $b$, and $c$ are integers. Expanding $\beta$ in powers of $g$, we have

@@

\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle = \sum_{m_1 = 1}^\infty \cdots \sum_{m_a = 1}^\infty \frac{(-\eta)^{m_1 + \dots + m_a}}{(1-\eta)^{m_1 + \dots + m_a}} \langle g^{m_1 + \dots + m_a} h_\alpha^b h_{\alpha \alpha}^c \rangle.

@@

Writing out the ensemble averaged quantity on the right hand side and defining $M \equiv m_1 + \dots + m_a$ yields

@@

\langle g^{m_1 + \dots + m_a} h_\alpha^b h_{\alpha \alpha}^c \rangle &=& \langle (\x^t \C^{-1} \x)^M (\x^t \C^{-1} \Q^\alpha \C^{-1} \x)^b (\x^t \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \C^{-1}\x )^c \rangle \nonumber \\

@@

\langle g^{m_1 + \dots + m_a} h_\alpha^b h_{\alpha \alpha}^c \rangle \approx N^{m_1 + \dots + m_a} \left[ \textrm{tr} (\C^{-1} \Q^\alpha )\right]^b (2 F_{\alpha \alpha})^c,

@@

\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle \approx \left[ \frac{\eta N}{1+\eta (N-1)} \right]^a \left[ \textrm{tr} (\C^{-1} \Q^\alpha )\right]^b (2 F_{\alpha \alpha})^c.

@@

With this result, taking the ensemble average of Eq. \eqref{eq:phatlossexpanded} is relatively straightforward. One simply replaces every copy of $\beta$ with $\beta_N \equiv \eta N / [1 + \eta (N-1)]$, every copy of $h_\alpha$ with $ \textrm{tr} (\C^{-1} \Q^\alpha )$, and every copy of $h_{\alpha \alpha}$ with $2F_{\alpha \alpha}$. Performing these substitutions and recalling \acl{make sure the reader will indeed be recalling at this point} that $\langle \hat{p}_\alpha \rangle = \langle h_\alpha \rangle / 2 F_{\alpha \alpha}$, we obtain

@@

\frac{\langle \hat{p}_\alpha^\textrm{loss} \rangle}{\langle \hat{p}_\alpha \rangle} = \frac{(1-\beta_N)^2}{1- 2\beta_N + \beta_N^2 \left[ \textrm{tr} (\C^{-1} \Q^\alpha) \right]^2 / F_{\alpha \alpha}}.

@@

\frac{\langle \hat{p}_\alpha^\textrm{loss} \rangle}{\langle \hat{p}_\alpha \rangle} \Bigg{|}_{\C \propto \I} = \frac{(1-\beta_N)^2}{(1-\beta_N)^2 + \beta_N^2} = \frac{(1-\eta)^2}{(1-\eta)^2 + N^2 \eta^2}.

@@

This is shown in Figure \ref{fig:analytic_sig_loss} for $N=20$. As expected, the signal loss factor is unity for $\eta=0$ (perfect estimation of $\C$) and goes to zero as $\eta$ approaches $1$.

>>>>

\Chat^{-1} = \frac{\C^{-1}}{1-\eta} \left[ \I - \frac{\eta \x \x^t \C^{-1}}{1+ \eta (g-1)}\right],

@@

\hat{F}_{\alpha \alpha} = \frac{F_{\alpha \alpha}}{(1-\eta)^2} \left[ 1 -\frac{\eta }{1+ \eta (g-1)} \frac{h_{\alpha \alpha}}{F_{\alpha \alpha}} + \frac{1}{2} \left( \frac{\eta }{1+ \eta (g-1)} \right)^2 \frac{h_\alpha^2}{F_{\alpha \alpha}}\right],

@@

\hat{p}_\alpha^\textrm{loss} &=& \frac{1}{2} \frac{h_\alpha}{F_{\alpha \alpha}} \left[ 1 - \frac{\eta g}{1+ \eta (g-1)}\right]^2  \left[ 1 -\frac{\eta }{1+ \eta (g-1)} \frac{h_{\alpha \alpha}}{F_{\alpha \alpha}} + \frac{1}{2} \left( \frac{\eta }{1+ \eta (g-1)} \right)^2 \frac{h_\alpha^2}{F_{\alpha \alpha}}\right]^{-1} \nonumber \\
&=& \frac{1}{2} \frac{h_\alpha}{F_{\alpha \alpha}}\left(1-2\beta + \beta^2\right)\sum_{n=0}^\infty \sum_{k=0}^n {n \choose k} \left( \frac{\varepsilon h_{\alpha \alpha}}{F_{\alpha \alpha}}\right)^k \left( - \frac{\varepsilon^2 h_\alpha^2}{2 F_{\alpha \alpha}}\right)^{n-k},

@@

where $\beta \equiv \eta g/[1+ \eta (g-1)]$ and $\varepsilon \equiv \beta / g$. To proceed, we take the ensemble average of $\hat{p}_\alpha^\textrm{loss}$ to understand its average statistical properties. Inspecting Eq. \eqref{eq:phatlossexpanded}, we see that its expectation value is a linear sum of terms of the form $\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \rangle$, where $a$, $b$, and $c$ are integers. Expanding $\beta$ in powers of $g$, we have

@@

\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \varepsilon^d \rangle = \sum_{m_1 = 1}^\infty \cdots \sum_{m_{a+d} = 1}^\infty \frac{(-\eta)^{m_1 + \dots + m_{a+d}}}{(1-\eta)^{m_1 + \dots + m_a}} \langle g^{m_1 + \dots + m_{a+d}-d} h_\alpha^b h_{\alpha \alpha}^c \rangle.

@@

Writing out the ensemble averaged quantity on the right hand side and defining $M \equiv m_1 + \dots + m_{a+d}$ yields

@@

\langle g^{m_1 + \dots + m_{a+d}} h_\alpha^b h_{\alpha \alpha}^c \rangle &=& \langle (\x^t \C^{-1} \x)^M (\x^t \C^{-1} \Q^\alpha \C^{-1} \x)^b (\x^t \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \C^{-1}\x )^c \rangle \nonumber \\

@@

\langle g^{m_1 + \dots + m_{a+d}} h_\alpha^b h_{\alpha \alpha}^c \rangle \approx N^{m_1 + \dots + m_{a+d}} \left[ \textrm{tr} (\C^{-1} \Q^\alpha )\right]^b (2 F_{\alpha \alpha})^c,

@@

\langle \beta^a h_\alpha^b h_{\alpha \alpha}^c \varepsilon^d \rangle \approx \left[ \frac{\eta N}{1+\eta (N-1)} \right]^a \left[ \frac{\eta }{1+\eta (N-1)} \right]^d \left[ \textrm{tr} (\C^{-1} \Q^\alpha )\right]^b (2 F_{\alpha \alpha})^c.

@@

With this result, taking the ensemble average of Eq. \eqref{eq:phatlossexpanded} is relatively straightforward. One simply replaces every copy of $\beta$ with $\beta_N \equiv \eta N / [1 + \eta (N-1)]$, every copy of $\varepsilon$ with $\varepsilon_N \equiv  \eta  / [1 + \eta (N-1)]$, every copy of $h_\alpha$ with $ \textrm{tr} (\C^{-1} \Q^\alpha )$, and every copy of $h_{\alpha \alpha}$ with $2F_{\alpha \alpha}$. Performing these substitutions and recalling \acl{make sure the reader will indeed be recalling at this point} that $\langle \hat{p}_\alpha \rangle = \langle h_\alpha \rangle / 2 F_{\alpha \alpha}$, we obtain

@@

\frac{\langle \hat{p}_\alpha \rangle}{\langle \hat{p}_\alpha^\textrm{loss} \rangle} = \frac{1- 2\varepsilon_N + \varepsilon_N^2 \left[ \textrm{tr} (\C^{-1} \Q^\alpha) \right]^2 / 2F_{\alpha \alpha}}{(1-\beta_N)^2}.

@@

\frac{\langle \hat{p}_\alpha \rangle}{\langle \hat{p}_\alpha^\textrm{loss} \rangle}  \Bigg{|}_{\C \propto \I} = \left( \frac{1-\varepsilon_N}{1-\beta_N} \right)^2 = \left[\frac{1+ \eta(N-2)}{1- \eta}\right]^2.

@@

This is shown in Figure \ref{fig:analytic_sig_loss} for $N=20$. As expected, the signal loss factor is unity for $\eta=0$ (perfect estimation of $\C$) and goes to infinity as $\eta$ approaches $1$.
